### 2019-08-06 ms/JSS/softwarereview.tex
setup for pasting code across now

\documentclass[softwarereview]{jss}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% declarations for jss.cls %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% reviewer
\Reviewer{Christopher J. Lortie\\NCEAS}
\Plainreviewer{Christopher Lortie}

%% about the software
\Softwaretitle{\pkg{Meta} 4.9-5}
\Softwaretitle{\pkg{Metafor} 2.1-0}
\Plaintitle{Meta 4.9-5}
\Plaintitle{Meta 4.9-5 and Metafor 2.1-0}

%% if different from \Softwaretitle also set
%% \Shorttitle{Aabel 1.5.7}
\Publisher{CRAN}
\Pubaddress{Vienna, Austria}
\Price{USD~0}
\URL{https://cran.r-project.org/web/packages/metafor/index.html}

%% publication information
%% NOTE: Typically, this can be left commented and will be filled out by the technical editor
%% \Volume{50}
%% \Issue{1}
%% \Month{June}
%% \Year{2012}
%% \Submitdate{2012-06-04}

%% address of reviewer
\Address{
  Christopher J. Lortie\\
  University of California Santa Barbara\\
  The National Center for Ecological Analysis and Synthesis\\
  Santa Barbara, CA, USA\\
  E-mail: \email{chris@christopherlortie.info}\\
  URL: \url{http://www.christopherlortie.info}
}

%% end of declarations %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%% include the review as usual
%% Note that you should use the \pkg{}, \proglang{} and \code{} commands.

\end{document}