### 2020-03-30 contrast.Rmd
getting there.

---
title: ""
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### A contrast of meta and metafor
Contrasted meta and metafor using the classic [Cochrane bronchoconstriction dataset]((https://www.cochrane.org/CD001183/AIRWAYS_nedocromil-sodium-for-preventing-exercise-induced-bronchoconstriction)) used in many texts to demonstrate meta-analsysis. [Here is a link to these data as a tidy csv files on figshare.](https://figshare.com/articles/Meta_thinking_data_preventing_exercise-induced_bronchoconstriction/7767020)

#### Meta
```{r meta, warning=FALSE, message=FALSE}
#load primary libraries
library(tidyverse) #tidy functions to read and format data
library(meta) #meta package current version

#Classic bronchoconstriction data from Cochrane
data <- read_csv(url("https://ndownloader.figshare.com/files/14460386"))

#continuous data
m1 <- metacont(Ne, Me, Se, Nc, Mc, Sc, data = data, sm = "SMD")
summary(m1)
forest(m1, layout = "JAMA")
metabias(m1)
funnel(m1)
radial(m1)

#cumulative evidence
forest(metacum(m1, pooled = "random"))

```

### Metafor
```{r, metafor, warning=FALSE, message=FALSE}
#load primary libraries
library(tidyverse) #tidy functions to read and format data
library(metafor) #metafor package current version

#Classic bronchoconstriction data from Cochrane
data <- read_csv(url("https://ndownloader.figshare.com/files/14460386"))

```