### 2020-03-30 ms/2020/appendix_a.Rmd

---
title: "Appendix A"
output: pdf_document
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

A list of objects included in each package when loaded into the R environment.  
```{r appendix, message=FALSE, warning=FALSE, echo=FALSE}
library(tidyverse)
library(kableExtra)

appendix <- read_csv("appendix.csv")
knitr::kable(appendix, "latex", booktabs = TRUE, longtable = TRUE) %>%
  column_spec(1, width = "2cm") %>%
  column_spec(2, width = "4cm") %>%
  column_spec(3, width = "5cm") %>%
  column_spec(5, width = "2cm") %>%
  kable_styling(latex_options = c("hold_position", "repeat_header"))
```