### 2020-04-01 downloads.Rmd
r package downloads completed

---
title: "downloads"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Package use
### Data
Grab the data from CRAN.

```{r data, warning=FALSE, message=FALSE}
library(tidyverse)
library(dlstats)
library(lubridate)

data <- cran_stats(c("meta", "metafor"))

#tidy up
data <- data %>%
  mutate(int = interval(ymd(start), ymd(end)), days = time_length(int, "day")) %>%
  select(days, everything(), -int)

#summary by net totals
totals <- data %>%
  group_by(package) %>% 
  summarise(total_downloads = sum(downloads), total_days = sum(days))

#summary by months
months <- data %>%
  mutate(month = month(start)) %>%
  group_by(month, package) %>% 
  summarise(total_downloads = sum(downloads), total_days = sum(days))

#write_csv(data, "data/downloads.csv")
#write_csv(totals, "data/total_downloads.csv")

```

### Viz
Check popularity.

```{r viz, warning=FALSE, message=FALSE}
ggplot(data, aes(package, downloads)) +
  geom_boxplot() +
  scale_color_brewer(palette = "Set1")

ggplot(months, aes(month, total_downloads, color = package)) +
  geom_point() +
  geom_smooth() +
  scale_color_brewer(palette = "Set1") +
  scale_x_continuous(limits = c(1, 12), breaks = seq(1,12))

```